---
title: "Bayesian exponential random graph modelling of KWL Burial network"
author:
  - Author One
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---


<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)
```

# Introduction

# Background

# Methods

## ERGM model

Exponential Random Graph Models (ERGMs) are a family of statistical model for network analysis to 
describe the formation of ties in a network. In traditional statistical model for network, a dyad, the pair of nodes, is measured by the presence of a tie according to the attributes of the nodes or ties, where each predictor variable for forming a tie is measured separately. In contrast to traditional model where variables not interacting with each other, ERGMs assumes the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes and ties (Robins et al. 2007). In a network using ERGMs, predictor variables, called "network statistics", are direct functions of the ties themselves that the probability of occurrence of the configuration of ties can be hypothesized (Morries et al. 2008). This approach enables networks to be constructed in a variety of assumption and contexts to model social network in the past using archaeological data (Amati et al. 2019). 

some concepts:
- endogenous statistics that ties are modelled based on the existence of other ties
- exogenous statistics that ties are formed according to monadic or dyadic node attributes

toturial: 
https://www.r-bloggers.com/ergm-tutorial/

## statistics
We want to know 

ERGM- consider transitivity by taking into account high-order k-triangles. Triangles and higher order cycles

We used ergm package in R for the model specification. Every term in an ERGM must have an associated algorithm for computing its value for network (Morris et al. 2008). The terms we selected for nwtwork specification include:

1. edges: ties, a measure of density, equal to kstar(1) for undirected networks

2. gwesp (geometrically weighted edgewise shared partners): a dyad-based configurations (not node-based) that counts the number of times that both nodes have a tie to a third node. It is a measure of the tendency for two actors who are tied to have x partners in common.

3. gwdegree (geometrically weighted degree distribution): a parametric form that represents the frequency distribution for nodal degree, each node counts only once. The local configuration is degrees and stars that are equivalent representations for the distribution of node-based edge counts

# Results

```{r model-ergm-1}
model.ergm <- burial_network_pre ~ 
  edges + # ties, a measure of density, equal to kstar(1) for undirected networks
  density +
  gwdegree(0.5, fixed = TRUE)  +
  triangle # triad relation, a measure of clustering or cohesion, also called transitive triple in undirected network
summary(model.ergm)
```

```{r model-ergm-2}
model.2 <- burial_network_pre ~ edges + # density
  gwesp(0.2, fixed = TRUE) +  # transitivity(cohesion; triangle), a tendency for those with shared partners to become tied, or tendency of ties to cluster together
  # triad-based clustering according to dyad-based configurations instead of node-based
  # number means weight parameter alpha, which controls the rate of declining marginal returns
  # fixed = TRUE means the scale parameter lambda is fit as a curved exponential-family model
  # not much difference in a range of 0-1.5, the lower the value of the scaling parameter, the less likely the model is to be degenerate
  # ergm can estimate the parameter from the data by using fixed=FALSE
  # http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
  gwdegree(0.8, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees
  # tendency of being in contact with multiple partners, measures of centralisation
  # distribution of node-based edge counts, each node counts only once
  # number means weight parameter decay
  # Morris et al. (2008)
summary(model.2)
```


```{r model-ergm-3}
model.3 <- burial_network_pre ~ edges +  # the overall density of the network
  nodematch('quantity') +    # quantity-based homophily, categorical nodal attribute
  # the similarity of connected nodes
  gwesp(0.2, fixed = TRUE) +    # transitivity
  gwdegree(0.8, fixed = TRUE)   # popularity
summary(model.3)
```


```{r model-bergm}
# Specify a prior distribution: normal distribution (low density and high transitivity)
prior.mean <- c(-3, 0, 1, 0) # prior mean corresponds to mean for each parameter
# follow Alberto Caimo et al. (2015) hospital example
prior.sigma <- diag(3, 4, 4) # covariance matrix structure
# normal distribution ðœƒ âˆ¼ Nd (ðœ‡prior , Î£prior ) as a suitable prior model for the model parameters of interests
# where the dimension d corresponds to the number of parameters, ðœ‡ is mean vector and Î£prior is a d Ã— d covariance matrix.

# Estimated posterior means, medians and 95% credible intervals for Models.3
# bergmM: Bayesian exponential random graphs models under missing data using the approximate exchange algorithm
parpost <- bergm(model.3,
                  prior.mean  = prior.mean,
                  prior.sigma = prior.sigma,
                  burn.in     = 200, # burn-in iterations for every chain of the populationm, drops the first 200
                  main.iters  = 2000, # iterations for every chain of the population
                  aux.iters   = 10000, # MCMC steps used for network simulation
                  nchains     = 8, # number of chains of the population MCMC
                  gamma       = 0.7) # scalar; parallel adaptive direction sampling move factor, acceptance rate

summary(parpost) 
```

The summary presents the statistics that each Î¸ corresponds to the parameter specified in ERGM previously. In general, positive mean indicates positive correlation, while negative mean indicates negative correlation. 

Î¸1 = number of ties
Î¸2 = individuals with the same abundance of burial goods
Î¸3 = gwesp is negative that rejects the assumption that actors with multiple partners in common are more likely to be directed connected
Î¸4 = 

# some exploration
1. there is different statistics between weighted and unweighed ties

Here is an example of inline code `r x` in the middle of a sentence. 

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
